---
title: "Process Predictors"
author: "Mikko Jimenez and Caitlin Mothes"
date: "`r Sys.Date()`"
---

# Process environmental predictors

As part of a broader pipeline for running a niche model in urban settings, this script is designed to provide a flexible process for reading in predictor variables, setting them to a consistent coordinate reference system, applying the desired processing analyses (if applicable) and cropping them to an extent of interest.

## Set up environment

```{r}
source("setup.R") 

# read in spatial layer that includes full extent of interest, in this case a boundary .shp
aoi <- read_sf("data/nature_in_the_city/gis/ft_collins_GMA_boundary.shp")

# list of paths to predictors (shapefiles, .tifs, etc.) separated by processing type
path_perccov <- c("data/input_raw/Impervious_Surface/impervious_surfaces_fort_collins_2015.shp",
                  "data/input_raw/Natural_Habitat/bc8f7f07-7560-4191-aa2e-8f274a1c29e7202046-1-133oxp5.v8wfi.shp",
                  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_TreeCanopy.tif",
                  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_Grass_Shrubs.tif",
                  "data/input_raw/Green_Space/nic_green_space_2015.shp"
                  )

path_dist <- c("data/input_raw/Hydrology/Hydrology.shp",
               "data/input_raw/Buildings/Buildings.shp",
               "data/input_raw/Traffic-Density/foCO_traffic.shp")

path_nodist <- c("data/input_raw/Traffic-Density/foCO_traffic.tif")


# create buffered_aoi
## we're in NAD83, so res and buffer are in meters
#buffered_aoi <- st_buffer(aoi, dist = 500)

# set desired resolution
resolution <- 100

# set a buffer for processing layers
buffer <- 3

# set output directory
output_path <- "data/input_processed"
```

## Execute `process_preds()` function for all variables

```{r}
start <- Sys.time()

# process traffic aadt separately
pred_aadt <- process_preds(
  "data/input_raw/Traffic-Density/foCO_traffic.shp",
  aoi,
  resolution = 100,
  type = "raw",
  field = "aadt",
  save = TRUE,
  output_path = output_path
)

# process perccov predictors
pred_perccov <- purrr::map(
  path_perccov,
  process_preds,
  aoi,
  resolution,
  type = "perccov",
  buffer,
  field = NULL,
  save = TRUE,
  output_path = output_path
) %>% set_names(tools::file_path_sans_ext(basename(path_perccov)))

# process distance predictors
pred_dist <- purrr::map(
  path_dist,
  process_preds,
  aoi,
  resolution,
  type = "dist",
  save = TRUE,
  output_path = output_path
) %>% set_names(tools::file_path_sans_ext(basename(path_dist)))

# process non distance predictors
pred_nodist <- purrr::map(
  path_nodist,
  process_preds,
  aoi,
  resolution,
  type = "nodist",
  save = TRUE,
  output_path = output_path
) %>% set_names(tools::file_path_sans_ext(basename(path_nodist)))

end <- Sys.time()


end-start
```

## Stack and save all predictors

```{r}
# stack rasters we want
combined_rast <- c(pred_aadt, rast(pred_perccov), rast(pred_dist))

# edit some of the names
# fix long name for natural habitat
names(combined_rast)[names(combined_rast) == "bc8f7f07-7560-4191-aa2e-8f274a1c29e7202046-1-133oxp5.v8wfi"] <- "Natural_habitat"

names(combined_rast[[1]]) <- "foco_aadt"

# Save the predictor raster stack
writeRaster(combined_rast, filename = file.path(output_path, "pred_stack.tif"), overwrite = TRUE)
```
