---
title: "Process Predictors"
author: "Mikko Jimenez and Caitlin Mothes"
date: "`r Sys.Date()`"
---

# Process environmental predictors

As part of a broader pipeline for running a niche model in urban settings, this script is designed to provide a flexible process for reading in predictor variables, setting them to a consistent coordinate reference system, applying the desired processing analyses (if applicable) and cropping them to an extent of interest.

```{r}
source("setup.R")

# AOI
aoi <- read_sf("data/ft_collins_GMA.shp")
```

# separate tifs

```{r}
# create a list of rasters with their corresponding lookup tables and output directories
raster_configs <- list(
  list(
    raster_path = "data/input_raw/Jasmine/Fort Collins IGS 2023- Reclassified Raster FGS Removed/IGS_LC_FOCO_2023.tif",
    lookup_table = data.frame(
      category_value = c(0, 1),
      land_cover = c("neg_space", "inf_greenspace")
    ),
    output_dir = "data/input_raw/Jasmine/Jasmine_inf_greenspace_sep"
  ),
  list(
    raster_path = "data/input_raw/landcover_mchale/FC_landcover.tif",
    lookup_table = data.frame(
      category_value = c(1, 2, 3, 4, 5, 6, 7),
      land_cover = c("TreeCanopy", "Grass_Shrubs", "BareSoil", "Water", "Buildings", "Roads_Railroads", "OtherPaved")
    ),
    output_dir = "data/input_raw/landcover_mchale/FC_landcover_sep"
  ),
   list(
    raster_path = "data/input_raw/Jasmine/Fort Collins IGS 2023- Full Land Cover Classes/FGS_RM_FOCO_LC_2023.tif",
    lookup_table = data.frame(
      category_value = c(1, 2, 3, 4, 5, 6),
      land_cover = c("Water", "Trees_Shrubs", "Low_Vegetation", "Barren", "Imperv_Surface", "Imperv_Roads")
    ),
    output_dir = "data/input_raw/Jasmine/Jasmine_lancover_sep"
   )
)

# loop each configuration and run the function
for (config in raster_configs) {
  tif_sep(config$raster_path, config$lookup_table, config$output_dir)
}

```

Combine impervious surface and impervious roads from Jasmine data

```{r}
is <- rast("data/input_raw/Jasmine/Jasmine_lancover_sep/lc_Imperv_Surface.tif")

isr <-  rast("data/input_raw/Jasmine/Jasmine_lancover_sep/lc_Imperv_Roads.tif")

is_all <- terra::merge(is, isr)

writeRaster(is_all, "data/input_raw/Jasmine/Jasmine_lancover_sep/lc_Imperv_all.tif", overwrite = TRUE)
```

Pull in wetland layer to replace hydrology

```{r}
library(arcpullr)

#watersheds <- get_spatial_layer("https://cnhp.colostate.edu/arcgis/rest/services/Wetland_Inv/Wetlands/MapServer/0")
aoi_wetlands <- aoi %>% 
  st_buffer(1000) %>% 
  st_transform(3857)

wetlands <- get_layer_by_envelope("https://cnhp.colostate.edu/arcgis/rest/services/Wetland_Inv/Wetlands/MapServer/0", aoi_wetlands)

# save to hydrology folder
st_write(wetlands, "data/input_raw/Hydrology/cnhp_foco_wetlands.shp")
```

## Set up environment

```{r}
# read in spatial layer that includes full extent of interest, in this case a boundary .shp
aoi <- read_sf("data/nature_in_the_city/gis/ft_collins_GMA_boundary.shp")

# list of paths to predictors (shapefiles, .tifs, etc.) separated by processing type
path_perccov <- c(
  # jasmine layers
  "data/input_raw/Jasmine/Jasmine_lancover_sep/lc_Imperv_Surface.tif",
  "data/input_raw/Jasmine/Jasmine_lancover_sep/lc_Trees_Shrubs.tif",
  "data/input_raw/Jasmine/Jasmine_lancover_sep/lc_Low_Vegetation.tif",
  "data/input_raw/Jasmine/Jasmine_inf_greenspace_sep/lc_inf_greenspace.tif",
  # mchale layers
  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_TreeCanopy.tif",
  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_Grass_Shrubs.tif",
  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_BareSoil.tif",
  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_OtherPaved.tif",
  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_Roads_Railroads.tif",
  # natural habitat city layer
  "data/input_raw/Natural_Habitat/bc8f7f07-7560-4191-aa2e-8f274a1c29e7202046-1-133oxp5.v8wfi.shp"
)

path_dist <- c("data/input_raw/Hydrology/cnhp_foco_wetlands.shp",
               "data/input_raw/Buildings/Buildings.shp",
               "data/input_raw/Jasmine/Jasmine_lancover_sep/lc_Imperv_Roads.tif")
               #"data/input_raw/Traffic-Density/foCO_traffic.shp")

#path_nodist <- c("data/input_raw/Traffic-Density/foCO_traffic.tif")


# create buffered_aoi
## we're in NAD83, so res and buffer are in meters
#buffered_aoi <- st_buffer(aoi, dist = 500)

# set desired resolution
resolution <- 100

# set a buffer for processing layers
buffer <- 3

# set output directory
output_path <- "data/input_processed"
```

## Execute `process_preds()` function for all variables

```{r}
start <- Sys.time()

# process traffic aadt separately
# pred_aadt <- process_preds(
#   "data/input_raw/Traffic-Density/foCO_traffic.shp",
#   aoi,
#   resolution = 100,
#   type = "raw",
#   field = "aadt",
#   save = TRUE,
#   output_path = output_path
# )

# process perccov predictors
pred_perccov <- purrr::map(
  path_perccov,
  process_preds,
  aoi,
  resolution,
  type = "perccov",
  buffer,
  field = NULL,
  save = TRUE,
  output_path = output_path
) %>% set_names(tools::file_path_sans_ext(basename(path_perccov)))

# process distance predictors
pred_dist <- purrr::map(
  path_dist,
  process_preds,
  aoi,
  resolution,
  type = "dist",
  save = TRUE,
  output_path = output_path
) %>% set_names(tools::file_path_sans_ext(basename(path_dist)))

# process non distance predictors
# pred_nodist <- purrr::map(
#   path_nodist,
#   process_preds,
#   aoi,
#   resolution,
#   type = "nodist",
#   save = TRUE,
#   output_path = output_path
# ) %>% set_names(tools::file_path_sans_ext(basename(path_nodist)))

end <- Sys.time()


end-start
```

## Stack and save all predictors

```{r}
# stack rasters we want
combined_rast <- c(rast(pred_perccov), rast(pred_dist))

# edit some of the names
# fix long name for natural habitat
names(combined_rast)[names(combined_rast) == "bc8f7f07-7560-4191-aa2e-8f274a1c29e7202046-1-133oxp5.v8wfi"] <- "Natural_habitat"

#names(combined_rast[[1]]) <- "foco_aadt"

# Save the predictor raster stack
writeRaster(combined_rast, filename = file.path(output_path, "pred_stack.tif"), overwrite = TRUE)
```
