# Run Connectivity Models in Julia

This module runs Omniscape in Julia. Omniscape is a newer model, based upon Circuitscape but better designed for spatially continuous input surfaces like habitat suitability models.

Omniscape User Guide: <https://docs.circuitscape.org/Omniscape.jl/latest/usage/#User-Guide> and GitHub: <https://github.com/Circuitscape/Omniscape.jl>

This section is run directly from Julia, as there are current incompatibilities with the newest version of Julia (1.11) and the `JuliaCall` R package. Future versions will hopefully be able to execute directly from this .Rmd. In the future, instructions for running directly from R can be found here: <https://github.com/Circuitscape/Circuitscape.jl/issues/381>.

## Install/Setup Julia

1.  Install Julia: <https://julialang.org/downloads/>

2.  Install Omniscape:

    ```         
    julia> using Pkg
    julia> Pkg.add("Omniscape")
    ```

    *Note: On my PC I had to add [this environmental variable](https://discourse.julialang.org/t/cannot-instantiate-env-in-julia-1-10-adwaita-icon-theme-jll-git-tree-hash-error-anyone-else/108105/5) in order for Circuitscape/Omniscape to install correctly:* JULIA_PKG_IGNORE_HASHES=true

    *Note: On Mac had to run this in terminal to load Julia after installation:*

      `. /Users/caitlinmothes/.zshrc`

3.  Check that all tests are passing:

    ```         
    julia> Pkg.test("Omniscape")
    ```

## Create Resistance files

```{r}
source("setup.R")
```

Omniscape takes in resistance files and prefers no values are less than one (in regards to its method for calculating sources from the resistance file).

Need to read in suitability rasters, multiply values by 10 (currently range 0-1) and then calculate the inverse and save resistance layers

```{r}
# read in resistance files
resistance_files <- map(list.dirs("data/output_sdm", recursive = FALSE), ~list.files(.x, pattern = ".tif", full.names = TRUE))

# calculate and save each resistance layer
walk(resistance_files, function(x, output_path = "data/output_resistance"){
  
  file <- terra::rast(x)
  
  # raster math
  file_convert <- 10 - file * 10 + 1
  
  # save
  if(!dir.exists(output_path)) {
    dir.create(output_path)
  }
  
  ## get file name
  file_name <- tools::file_path_sans_ext(basename(x))
  writeRaster(file_convert, paste0(output_path, "/", file_name, "_resistance.tif"))
  
})

```

## Create ini files

Details on Omniscape settings can be found here: <https://docs.circuitscape.org/Omniscape.jl/stable/usage/#Running-Omniscape>

In R, create .ini file for each species

```{r}
# function to create ini paths
create_ini <- function(resistance_file,
                       radius = 100, # check on this, this is # pixels
                       block_size = 1,
                       project_name,
                       source_from_resistance = TRUE,
                       r_cutoff = 50,
                       calc_normalized_current = TRUE,
                       parallelize = TRUE,
                       parallel_batch_size = 20,
                       solver = "cholmod",
                       write_raw_currmap = TRUE,
                       species,
                       ini_path = "ini/") {
  
  config <- list(
    Required = list(
      resistance_file = resistance_file,
      radius = radius,
      block_size = block_size,
      project_name = project_name
    ),
    General_options = list(
      source_from_resistance = source_from_resistance,
      r_cutoff = r_cutoff,
      calc_normalized_current = calc_normalized_current,
      parallelize = parallelize,
      parallel_batch_size = parallel_batch_size,
      solver = solver
    ),
    Outpt_options = list(
      write_raw_currmap = write_raw_currmap
    )
  )
  
  if(!dir.exists(ini_path)){
    dir.create(ini_path)
  }
  
  write.ini(config, paste0(ini_path, "/", species, ".ini"))
  
  cat(paste("INI file for", species, "created!"))
  
}
```

```{r}
# vectors to map over

```

## Run Omniscape

Set number of threads to use **(from Command Prompt)**

```         
julia -t 4
```

Run Omniscape from the .ini file

```         
using Omniscape
run_omniscape("Desktop/dsri_foco_habicon/ini/testing.ini")
```

## Explore Outputs

```{r}
cum_curr_map <- rast("data/output_connectivity/Agelaius_phoeniceus/cum_currmap.tif")

norm_cum_currmap <- rast("data/output_connectivity/Agelaius_phoeniceus/normalized_cum_currmap.tif")
```

```{r}
tmap_mode("view")

tm_shape(cum_curr_map) +
  tm_raster(style = "cont") +
  tm_shape(norm_cum_currmap) +
  tm_raster(style = "cont")
```
