# Run Connectivity Models in Julia

This module runs Omniscape in Julia. Omniscape is a newer model, based upon Circuitscape but better designed for spatially continuous input surfaces like habitat suitability models.

Omniscape User Guide: <https://docs.circuitscape.org/Omniscape.jl/latest/usage/#User-Guide> and GitHub: <https://github.com/Circuitscape/Omniscape.jl>

This workflow executes Omniscape in Julia, and as there are current incompatibilities with the newest version of Julia (1.11) and the `JuliaCall` R package, must be run externally. Future versions will hopefully be able to execute directly from this .Rmd. In the future, instructions for running directly from R can be found here: <https://github.com/Circuitscape/Circuitscape.jl/issues/381>.

## Install/Setup Julia and Omniscape

1.  Install Julia: <https://julialang.org/downloads/>

2.  Install Omniscape:

    ```         
    julia> using Pkg
    julia> Pkg.add("Omniscape")
    ```

    *Note: On PC I had to add [this environmental variable](https://discourse.julialang.org/t/cannot-instantiate-env-in-julia-1-10-adwaita-icon-theme-jll-git-tree-hash-error-anyone-else/108105/5) in order for Circuitscape/Omniscape to install correctly:* JULIA_PKG_IGNORE_HASHES=true

    *Note: On Mac had to run this in terminal to load Julia after installation:*

      `. /Users/caitlinmothes/.zshrc`

3.  Check that all tests are passing:

    ```         
    julia> Pkg.test("Omniscape")
    ```

## Create Resistance files

```{r}
source("setup.R")
```

Omniscape takes in resistance files and prefers no values are less than one (in regards to its method for calculating sources from the resistance file).

Need to read in suitability rasters, multiply values by 10 (currently range 0-1) and then calculate the inverse and save resistance layers.

```{r}
# read in suitability files
suitability_files <- map(list.dirs("data/output_sdm", recursive = FALSE), ~list.files(.x, pattern = ".tif", full.names = TRUE))

# calculate and save each resistance layer
walk(suitability_files, function(x, output_path = "data/output_resistance"){
  
  file <- terra::rast(x)
  
  # raster math
  file_convert <- 10 - file * 10 + 1
  
  # save
  if(!dir.exists(output_path)) {
    dir.create(output_path)
  }
  
  ## get file name
  file_name <- tools::file_path_sans_ext(basename(x))
  writeRaster(file_convert, paste0(output_path, "/", file_name, "_resistance.tif"), overwrite = TRUE)
  
})

```

## Create ini files

Details on Omniscape settings can be found here: <https://docs.circuitscape.org/Omniscape.jl/stable/usage/#Running-Omniscape>

In R, create an .ini file for each species

```{r}
# function to create ini paths
create_ini <- function(resistance_file,
                       radius = 100, # check on this, this is # pixels
                       block_size = 1,
                       project_name,
                       source_from_resistance = TRUE,
                       r_cutoff = 5, #Using 5, which represents >50% prob of occ can act as a source
                       calc_normalized_current = TRUE,
                       parallelize = TRUE,
                       parallel_batch_size = 20,
                       solver = "cholmod",
                       write_raw_currmap = TRUE,
                       species,
                       ini_path = "ini/") {
  
  config <- list(
    Required = list(
      resistance_file = resistance_file,
      radius = radius,
      block_size = block_size,
      project_name = project_name
    ),
    General_options = list(
      source_from_resistance = source_from_resistance,
      r_cutoff = r_cutoff,
      calc_normalized_current = calc_normalized_current,
      parallelize = parallelize,
      parallel_batch_size = parallel_batch_size,
      solver = solver
    ),
    Outpt_options = list(
      write_raw_currmap = write_raw_currmap
    )
  )
  
  if(!dir.exists(ini_path)){
    dir.create(ini_path)
  }
  
  write.ini(config, paste0(ini_path, "/", species, ".ini"))
  
}
```

Execute function and create ini files

```{r}
# species to map over
species <- map(suitability_files, ~tools::file_path_sans_ext(basename(.x)) %>% str_remove("_prediction"))

# paths to resistance files (ensure same order as species vector)
resistance_files <-  list.files("data/output_resistance/", pattern = ".tif", full.names = TRUE) %>% normalizePath()

# execute ini function
walk2(resistance_files, species, ~create_ini(resistance_file = .x, species = .y,
                                             project_name = paste0(normalizePath(paste0("data/output_connectivity/")), "/", .y, "_", Sys.Date())))

```

## Run Omniscape

Start Julia and set number of threads to use **(from Command Prompt).**

```{bash}
julia -t 4
```

Execute Omniscape for all species

```{julia}
using Omniscape
using Glob  # For easier file pattern matching

# Define the folder containing the .ini files
folder_path = "/Users/caitlinmothes/Desktop/dsri_foco_habicon/ini"

# List all .ini files in the specified folder
ini_files = glob("*.ini", folder_path)

# Function to process each .ini file
function process_ini_files(files)
    for ini_file in files
        println("Processing: ", ini_file)
        try
            run_omniscape(ini_file)
        catch e
            println("Error processing file ", ini_file, ": ", e)
        end
    end
end

# Run the function on the list of .ini files
process_ini_files(ini_files)

```

## Inspect Maps

```{r}
cum_maps <- map(list.dirs("data/output_connectivity/", recursive = FALSE), ~terra::rast(paste0(.x, "/cum_currmap.tif")))

norm_maps <- map(list.dirs("data/output_connectivity/", recursive = FALSE), ~terra::rast(paste0(.x, "/normalized_cum_currmap.tif")))


# map them out
map(cum_maps, ~tm_shape(.x) + tm_raster(style = "cont"))

map(norm_maps, ~tm_shape(.x) + tm_raster(style = "cont"))
```
