# Run Connectivity Models in Julia

This module runs Omniscape in Julia. Omniscape is a newer model, based upon Circuitscape but better designed for spatially continuous input surfaces like habitat suitability models.

Omniscape User Guide: <https://docs.circuitscape.org/Omniscape.jl/latest/usage/#User-Guide> and GitHub: <https://github.com/Circuitscape/Omniscape.jl>

This section is run directly from Julia, as there are current incompatibilities with the newest version of Julia (1.11) and the `JuliaCall` R package. Future versions will hopefully be able to execute directly from this .Rmd. In the future, instructions for running directly from R can be found here: <https://github.com/Circuitscape/Circuitscape.jl/issues/381>.

## Install/Setup Julia

1.  Install Julia: <https://julialang.org/downloads/>

2.  Install Omniscape:

    ```         
    julia> using Pkg
    julia> Pkg.add("Omniscape")
    ```

    *Note: On my PC I had to add [this environmental variable](https://discourse.julialang.org/t/cannot-instantiate-env-in-julia-1-10-adwaita-icon-theme-jll-git-tree-hash-error-anyone-else/108105/5) in order for Circuitscape/Omniscape to install correctly:* JULIA_PKG_IGNORE_HASHES=true

3.  Check that all tests are passing:

    ```         
    julia> Pkg.test("Omniscape")
    ```

## Run Omniscape

Set number of threads to use **(from Command Prompt)**

```         
julia -t 4
```

Run Omniscape from the .ini file

```         
using Omniscape
run_omniscape("Desktop/dsri_foco_habicon/ini/testing.ini")
```

## Explore Outputs

```{r}
cum_curr_map <- rast("data/output_connectivity_1/cum_currmap.tif")

norm_cum_currmap <- rast("data/output_connectivity_1/normalized_cum_currmap.tif")
```

```{r}
tmap_mode("view")

tm_shape(cum_curr_map) +
  tm_raster(style = "cont") +
  tm_shape(norm_cum_currmap) +
  tm_raster(style = "cont")
```
