# Workflow to explore SDM results

## Setup

```{r}
source("setup.R")

# list all directories in output_sdm
output_dirs <- list.dirs("data/output_sdm", recursive = FALSE)

specs <- basename(output_dirs)
```

### Load in data

```{r}
walk(output_dirs, ~load(list.files(.x, pattern = ".RData", full.names = TRUE), envir = .GlobalEnv))
```

## Visualize results

### Model Performance

#### Tuning Results

Look at omission rates and validation AUC

```{r}
evalplot.stats(
  e = Agelaius_phoeniceus_SDM_output$all_mods,
  stats = c("auc.diff", "cbi.val", "or.10p"),
  color = "fc",
  x.var = "rm",
  error.bars = FALSE
  #dodge = 0.5
)
```

```{r}

```

### Null Models

```{r}
#testing
load("data/output_sdm_2025-02-28/Agelaius phoeniceus/Agelaius_phoeniceus_SDM_results.RData")
```

Compare empirical results to simulated. Here hoping for empirical eval metrics that are significantly better than a null model.

```{r}
null.emp.results(Agelaius_phoeniceus_SDM_output$null_mods) %>% DT::datatable()
```

Plots of the null model results as a histogram. In this example, the empirical validation CBI value (solid red line) is significantly higher than random, as it is higher than the 99th quantile of the null values (dashed purple line). For reference, this plot also includes the null 95th quantile (dashed blue line) and the null mean (solid blue line).

```{r}
evalplot.nulls(Agelaius_phoeniceus_SDM_output$null_mods, stats = c("or.10p", "auc.val", "cbi.val"), plot.type = "histogram")
```

### Variable Importance

```{r}
ggplot(Agelaius_phoeniceus_SDM_output$variable_importance, aes(x = reorder(Variable, -Permutation_importance), y = Permutation_importance)) +
  geom_col() +
  geom_errorbar(aes(ymin = Permutation_importance - sd, 
                    ymax = Permutation_importance + sd),
                width = 0.2)+
  xlab("")
```

### Response Curves

```{r}
# get list of variables
vars <- Agelaius_phoeniceus_SDM_output$variable_importance$Variable

rc <- Agelaius_phoeniceus_SDM_output$response_curves
```

```{r}
library(patchwork)

plots <- map(vars, ~ggplot(rc, aes(x = !!sym(.x), y = !!sym(paste0("preds_", .x)))) +
  geom_line() +
  xlab(.x) +
  ylab("Presence Probability"))

wrap_plots(plots, ncol = 3)
```
