---
title: "Code to process predictors for urban niche model"
author: "Mikko Jimenez, Josh Carrell, and Caitlin Mothes"
date: "'r Sys.Date()'"
output: html_notebook
---

## Process predictors

As part of a broader pipeline for running a niche model in urban settings, this script is designed to provide a flexible process for reading in predictor variables, setting them to a consistent coordinate reference system, and cropping them to an extent of interest. 

Note: it is currently unclear if predictors will be expected to be downloaded prior to running this script or if we will be pulling predictors directly from an API. I've written code for each situation below. 

## set up environment
```{r}
source("setup.R") 

# read in spatial layer that includes full extent of interest, in this case a boundary .shp
aoi <- read_sf("data/nature_in_the_city/gis/ft_collins_GMA_boundary.shp")

# list of paths to predictors (shapefiles, .tifs, etc.)
path_perccov <- c("data/input_raw/Impervious_Surface/impervious_surfaces_fort_collins_2015.shp",
                  "data/input_raw/Natural_Habitat/bc8f7f07-7560-4191-aa2e-8f274a1c29e7202046-1-133oxp5.v8wfi.shp",
                  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_TreeCanopy.tif",
                  "data/input_raw/landcover_mchale/FC_landcover_sep/lc_Grass_Shrubs.tif",
                  "data/input_raw/Green_Space/nic_green_space_2015.shp"
                  )
path_dist <- c("data/input_raw/Hydrology/Hydrology.shp",
               "data/input_raw/Buildings/Buildings.shp")
path_nodist <- c("data/input_raw/Traffic-Density/foCO_traffic.tif")

# we're in NAD83, so res and buffer are in meters
# create buffered_aoi
# add a dist argument
buffered_aoi <- st_buffer(aoi, dist = 500)
# set desired resolution
resolution <- 100
# set a buffer
buffer <- 3

# set output directory
output_path <- "data/input_processed"
```

## write function that reads in predictors and matches CRS and extent to the aoi
```{r}
# write function that reads in predictors and matches CRS and extent to the aoi
process_preds <- function(input_path, aoi_path, resolution, type, buffer, save, output_path) {
  
  # aoi buffer idea
  #buffered_aoi <- st_buffer(aoi, dist = dist)
  
  # Initialize variables to hold the processed objects
  rast_shapefile <- NULL
  raster_file <- NULL
  output_object <- NULL
  
  if (grepl("\\.shp$", predictor_path, ignore.case = TRUE)) {
    # Read in shapefile
    shapefile <- st_read(predictor_path)
    
    # Set CRS to match the aoi
    shapefile <- st_transform(shapefile, crs = crs(aoi))
    
    # Set extent to match the aoi (crop if necessary)
    shapefile <- st_crop(shapefile, st_bbox(aoi))
    
    ## Rasterize the shapefile
    # Create an empty raster aoi with the specified resolution
    raster_aoi <- rast(ext(aoi), resolution = resolution)
    # Rasterize shapefile 
    rast_shapefile <- terra::rasterize(shapefile, raster_aoi, fun = mean)
    # Set CRS to match the shapefile
    crs(rast_shapefile) <- st_crs(shapefile)$proj4string
    
    # Perform calculations based on 'type' argument
    if (type == "dist") {
      rast_shapefile <- terra::distance(rast_shapefile)
      
    } else if (type == "perccov") {
      if (is.null(buffer)) {
        stop("Buffer must be specified when calculating percent cover.")
      }
      # Calculate percent cover within the buffer distance
      rast_shapefile <- terra::focal(rast_shapefile, w = buffer, fun = function(x) mean(!is.na(x), na.rm = TRUE) * 100)
    }
    
    # Extract filename without extension for naming 
    file_name <- tools::file_path_sans_ext(basename(predictor_path))
    
    # Store the processed object with the filename
    output_object <- rast_shapefile
    
    names(output_object) <- file_name
    
  } else if (grepl("\\.tif$", predictor_path, ignore.case = TRUE)) {
    # Read in raster and aoi files
    raster_file <- rast(predictor_path)
    
    # Set aoi temporarily to raster CRS - so we don't need to project full, large rasters
    shapefile_temp <- st_transform(aoi, crs = crs(raster_file))
    
    # crop to the temp shapefile
    raster_file <- crop(raster_file, shapefile_temp)
    
    # Convert the aoi to a raster with the same extent and resolution
    aoi_raster <- rast(ext(aoi), resolution = resolution, crs = crs(aoi))
    
    # Project cropped raster into aoi CRS
    raster_file <- project(raster_file, aoi_raster)
    
    # Perform calculations based on 'type' argument
    if (type == "dist") {
      raster_file <- terra::distance(raster_file)
      
    } else if (type == "perccov") {
      if (is.null(buffer)) {
        stop("Buffer must be specified when calculating percent cover.")
      }
      # Calculate percent cover within the buffer distance
      raster_file <- terra::focal(raster_file, w = buffer, fun = function(x) mean(!is.na(x), na.rm = TRUE) * 100)
    }
    
    # Extract filename without extension for naming
    file_name <- tools::file_path_sans_ext(basename(predictor_path))
    
    # Store the processed object with the filename
    output_object <- raster_file
    
    names(output_object) <- file_name
    
  } else {
    stop("Unsupported file type.")
  }
  
  # Save the processed objects to the specified directory if 'save' is TRUE
  if (save) {
    if (!dir.exists(output_path)) {
      dir.create(output_path, recursive = TRUE)
    }
    
    if (!is.null(rast_shapefile)) {
      # Write raster file
      writeRaster(rast_shapefile, filename = file.path(output_path, paste0(file_name, ".tif")), overwrite = TRUE)
    }
    
    if (!is.null(raster_file)) {
      # Write raster file
      writeRaster(raster_file, filename = file.path(output_path, paste0(file_name, ".tif")), overwrite = TRUE)
    }
  }
  
  # Return the processed object with the filename
  return(output_object)
}
```

# process predictors in parallel 
```{r}
# process perccov predictors
pred_perccov <- purrr::map(path_perccov, process_preds, aoi, resolution, type="perccov", buffer, save=FALSE, output_path) %>% set_names(tools::file_path_sans_ext(basename(path_perccov)))

# process distance predictors
pred_dist <- purrr::map(path_dist, process_preds, aoi, resolution, type="dist", buffer, save=FALSE, output_path) %>% set_names(tools::file_path_sans_ext(basename(path_dist)))

# process non distance predictors
pred_nodist <- purrr::map(path_nodist, process_preds, aoi, resolution, type="nodist", buffer, save=FALSE, output_path) %>% set_names(tools::file_path_sans_ext(basename(path_nodist)))
```
# stack all the rasters and save
```{r}
# stack rasters
combined_rast <- terra::rast(c(pred_perccov, pred_dist, pred_nodist))
plot(combined_rast)

# Save the predictor raster stack
writeRaster(combined_rast, filename = file.path(output_path, "pred_stack.tif"), overwrite = TRUE)
```



